---
title: "lmds"
output: github_document
---

[![Build Status](https://travis-ci.org/dynverse/lmds.svg?branch=master)](https://travis-ci.org/dynverse/lmds) 


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%",
  fig.path = "man/figures/",
  message = FALSE,
  dpi = 300
)
set.seed(1)
library(tidyverse)
```

# `lmds`:  Landmark Multi-Dimensional Scaling

Landmark Multi-Dimensional Scaling (LMDS) is an extension of classical Torgerson MDS.
LMDS aims to produce the same dimensionality reduction as `cmdscale()`, but scale
linearly with respect to the number of samples.

```{r compare}
library(lmds)
x <- as.matrix(iris[,1:4])
dimred <- lmds(
  x,
  ndim = 2, 
  distance_method = "euclidean", 
  landmark_method = "sample",
  num_landmarks = 20
)
qplot(dimred[,1], dimred[,2]) + labs(title = "lmds()") + theme_classic()

dimred <- cmdscale(dist(x))
qplot(dimred[,1], dimred[,2]) + labs(title = "cmdscale()") + theme_classic()
```

The execution time of `lmds()` scales linearly with respect to the dataset size.
```{r timings, echo=FALSE}
largex <- Matrix::rsparsematrix(nrow = 100000, ncol = 10000, density = .01)

log <- list()
num_samples <- 100
time <- 0
while (time < 10) {
  subx <- largex[seq_len(num_samples),]
  time <- system.time({
    dimred <- lmds(subx)
  })[["user.self"]]
  log[[length(log) + 1]] <- tibble(method = "lmds()", num_samples, time)
  num_samples <- num_samples * 1.5
}

num_samples <- 100
time <- 0
while (time < 10) {
  subx <- largex[seq_len(num_samples),]
  time <- system.time({
    dist <- dist(subx)
    dimred <- cmdscale(dist)
  })[["user.self"]]
  log[[length(log) + 1]] <- tibble(method = "cmdscale()", num_samples, time)
  num_samples <- num_samples * 1.5
}

logdf <- bind_rows(log)

ggplot(logdf, aes(num_samples, time, colour = method)) + 
  geom_point() +
  geom_line() +
  theme_classic()
```

## Latest changes

Check out `news(package = "lmds")` or [NEWS.md](NEWS.md) for a full list of changes.

<!-- This section gets automatically generated from inst/NEWS.md, and also generates inst/NEWS -->

```{r news, echo=FALSE, results="asis"}
cat(dynutils::recent_news())
```
